---
title: "Phenotypic Linkage Network Demonstration"
author: "Cynthia Sandor & Viola Volpato"
date: 'Last update: `r date()`'
output:
  html_document:
    fig_height: 5
    fig_width: 20
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
      toc_depth: 5
  pdf_document:
    toc: yes
---

## Set up your work environnement

To evaluate a phenotypic linkage network you must:

1) Install some R/Bioconductor packages
2) Indicate a directory where you will download PLN data (require ~10Gb on your disk)
3) Indicate a list of genes (format gene symbol)

### The following packages must be installed and loaded

To install a general R package you must type: install.packages("gridExtra")
To install a Bioconductor R package you must type: BiocManager::install("biomaRt"). 
You must install before BiocManager: install.packages("BiocManager")

```{r load the required package,echo=FALSE}
library(biomaRt)
library(ggplot2)
library(gridExtra)
library(DT)
library(igraph)
library(ggnetwork)
library(network)
```

### Work directory

You must indicate a work directory where the PLN files will be downloaded

```{r work directory,echo=FALSE}
#dir_work<-"/Users/dpag0499/OneDrive - Cardiff University/Documents/PROJECTS/PLN_Demo_Workshop/test"
#dir_pln<-"/Users/dpag0499/OneDrive - Cardiff University/Documents/PROJECTS/PLN_Demo_Workshop/src/Rscripts/clustering_analysis"

dir_work<-"where_you_downloaded_PLN_files"
dir_pln<-"where_you_downloaded_PLN_Github"


```

You must download the following tar.gz including all the PLN file in your work directory:
[pln.tar.gz](https://drive.google.com/open?id=1wDFoldeFf8JjiWRhc2IjEtKhZ0nVtFwp)

```{r untar the tar.gz,echo=FALSE}
##untar(paste(dir_work,"/pln.tar","",sep=""))
## ADVICE FOR Windows Users!
```
### List of genes 

You must specify where your list of genes is

```{r list of genes,echo=FALSE}
f_gene<-paste(dir_work,"/pln/list_gene_symbol",sep="")
list_gene<-read.table(f_gene,h=F)
DT::datatable(list_gene, rownames = TRUE, caption = "List of genes", escape = FALSE, extensions = 'Buttons', options = list(dom = 'ftipr'))
```

### Set up the parameters

Here are the parameters required to evaluate the functional links between your list of genes by using PLN

```{r parameters,echo=FALSE}
parameters<-list()
parameters[["nb_gene"]]<-100
parameters[["nb_links"]]<-1000000
parameters[["f_size"]]<-paste(dir_work,"/pln/hs_68_cds_max_mean_length",sep="")
parameters[["f_pln"]]<-paste(dir_work,"/pln/wl_general.final.scale.ord",sep="")
parameters[["f_gene"]]<-f_gene
parameters[["nb_sim"]]<-100
parameters[["f_list_dataset"]]<-paste(dir_work,"/pln/list_file_rescore_wl_info_wl.general",sep="")
parameters[["dir_data"]]<-paste(dir_work,"/pln",sep="")
df_parameters<-data.frame("Parameter"=names(unlist(parameters)),"Value"=unlist(parameters))
DT::datatable(list_gene, rownames = TRUE, caption = "List of genes", escape = FALSE, extensions = 'Buttons', options = list(dom = 'ftipr'))
```

### Functional dataset used 

We used the final linkage network built with the following datasets

```{r list dataset,echo=FALSE}
  
  # get info links dataset
  info_dataset<-read.table(parameters[["f_list_dataset"]],h=F)
  colnames(info_dataset)<-c("File","Type")
  DT::datatable(info_dataset, rownames = TRUE, caption = "List of dataset", escape = FALSE, extensions = 'Buttons', options = list(dom = 'ftipr'))
  
```

## Functionnal Clustering

We run an Rscript to evaluate and represent the functional relationship between genes.
There are two parts displayed in two different plots:

Plot1: Histogramm

We evaluate whether a given set of genes demonstrates unusually similar functionality by examining the extent to which those genes cluster within the PLN. For this, we compare the sum of weighted links observed between
these genes and between an equal number of random genes and we compute an empirical p-value (shown in the plot). As the
coding-sequence (CDS) length and the network connectivity (degree) can bias functional network
associations, the randomized genes are matched on CDS length and on the degree to the original genes.


Plot2: Gene network representation

We extract the most contributing resources, among the integrated datasets to build a PLN, for any functional associations found; (3) represents the extracted subnetwork with links colored by the most contributing resources.


```{r run fonctional analysis,echo=FALSE}

setwd(dir_pln)
source("functional_clustering_analysis_p.R")
functional_clustering_analysis(parameters)
```

